Secure Composition of Robust and Optimising Compilers

### 2023-07-07 popl/main.tex
Shorten sec 6 up to blue lang

%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
\section{Case Study: Preliminary Definitions\pages{6}}\label{sec:casestud:defs}

@@

\subsection{Shared Language Definitions}\label{subsec:cs:defs}
\MPin{
	explain with example ? leave text for app
}

\begin{gather*}
  \begin{aligned}
  &\mi{(Finals)}~\finalexpr\bnfdef\ \valueexpr \mid x \hspace{0.5cm}
  \mi{(Values)}~\valueexpr\bnfdef\ n \hspace{0.5cm}
  \mi{(Types)}~\type\bnfdef\ \natt \\
  \mi{(Expressions)}~\expr&\bnfdef\ \finalexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{x}{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
  \mi{(Symbols)}~\function&\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \hspace{1.00cm}
    \text{where}\ \lbinop{}{}\in\left\{+,-,\times,<,/\right\}\\
  &\mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
  \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
  \end{aligned}
\end{gather*}

Above is the base syntax of all the programming languages of this paper.
\MK{this needs a complete rewrite}
Expressions ($\expr$) can only be evaluated in the form of runtime--terms ($\rtt{\statevar}{\expr}$).
States ($\statevar$) have a flow- ($\cfstate$) and a memory state ($\memstate$).
As for the flow state, the list of component--level names ($\commlib$), the list of symbols ($\library$), and a stack of continuations tagged with the name of a symbol ($\ectxstack$) are necessary for calling and returning while emitting the right transfer tag ($\sandboxtag$) depending on whether there is a context switch from context to component ($\comptoctx$), from component to context ($\ctxtocomp$), or whether control flow stays internal to either ($\nocomm$).
The memory state contains both a context--level ($\heap[^{\ctx}]$) and a component--level heap ($\heap[^{comp}]$) as well as a store ($\store$).
Effectively, it is syntactically impossible for a context--level runtime--term to have a pointer into the component--level heap and vice--versa.
Pointers are non-nullable and there is no syntactic way to let them point into the middle of an object.
However, pointers always point to arrays and the individual elements of that array are read- and writeable by means of adding an offset that may be negative.
Heap locations are second--class. %The syntax is constrained as such that one cannot simply guess a heap location.
Stores map identifiers to pointer metadata.
This metadata is a tuple with the location ($\loc$), the control--tag ($\sandboxtag$), poison--tag ($\poison$), and the size of the allocated object ($n$).
The control-tag allows to distinguish whether a location indexes into $\heap[^{\ctx}]$ or $\heap[^{\comp}]$.
The poison--tag marks a location as freed ($\poisoned$) without removing the entry from $\store$.
If one removes an entry, for example, by evaluating $\ldelete{x}$, then there is no way to distinguish between freed pointers and free variables.
The latter should render evaluation as stuck, while the former should emit a $\ev{Dealloc}\ \loc$ event regardless of wether the object a pointer points to is still allocated or not.
Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) have disjoint sets of symbols, one for context-level ($\library_{\ctx}$) and one for component-level symbols ($\library_{\comp}$).

\begin{gather*}
  \begin{aligned}
  \mi{(Eval.~Ctx.)}~\ectx&\bnfdef\ \hole{\cdot} \mid \lbinop{\ectx}{\expr} \mid \lbinop{\valueexpr}{\ectx} \mid \lget{x}{\ectx} \mid \lnew{x}{\ectx}{\expr} \mid \lset{x}{\ectx}{\expr}\\
    & \mid \lset{x}{\valueexpr}{\ectx} \mid \lreturn{\ectx} \mid \lcall{g}{\ectx} \mid \lifz{\ectx}{\expr[_{1}]}{\expr[_{2}]} \\
  \end{aligned}\\
  \begin{aligned}
  \mi{(States)}~\statevar&\bnfdef \tup{\cfstate}{\sandboxtag,\memstate} &
  \mi{(Poison)}~\poison&\bnfdef\ \poisoned \mid \poisonless &
  \mi{(Locations)}~\loc\in\nat \\
  \end{aligned}\\
  \begin{aligned}
  \mi{(Heaps)}~\heap&\bnfdef\ \hole{\cdot} \mid n,\heap &
  \mi{(Stores)}~\store&\bnfdef\ \hole{\cdot} \mid \storeel{x}{\loc}{\sandboxtag}{\poison}{n},\store &
  \mi{(Transfer~Tags)}~\comm\bnfdef\ \nocomm \mid\ \ctxtocomp \mid\ \comptoctx \\
  \end{aligned}\\
  \begin{aligned}
  \mi{(Runtime~Terms)}~\runtimetermvar&\bnfdef\rtt{\statevar}{\expr} \mid \lightning &
  \mi{(Component~Names)}~\commlib&\bnfdef\ \hole{\cdot} \mid g,\commlib \\
  \end{aligned}\\
  \begin{aligned}
  \mi{(Continuation~Stacks)}~&\ectxstack\bnfdef\ \hole{\cdot} \mid \tup{\ectx}{g},\ectxstack \\
  \end{aligned}
\end{gather*}
All programming languages presented in this section have an environmental semantics and all but $\src{L_{\tmssafe}}$ are dynamically typed.
Each evaluation may emit the empty event ($\emptyevent$), a program crash token ($\lightning$), or a pair containing a pre--event ($\event[_{b}]$) and the corresponding control--tag ($\sandboxtag$) indicating whether the event fired during an evaluation of a context or a component.
Note that the reflexive-transitive closure used in this paper behaves like a big-step semantics, where the reflexive case is ony for the empty trace and finals, and the emitted trace does not contain empty events.
Most rules are entirely standard and omitted.
Heaps may need to grow when allocating, which the section writes as $\growh{\heap}{n}$, denoting that $\heap$ grows by $n$ cells.
Indexing into a heap, looking up symbols in libraries, or accessing the metadata of a pointer from a store are all written as a function call.
For example, the value located at the fifth position in heap $\heap$ is written as $\heap(5)$.
The operation to set the fifth position in heap $\heap$ to value $\valueexpr$ is written as $\seth{\heap}{5}{\valueexpr}$ and yields a new heap that is entirely the same to $\heap$ modulo the fifth position.

@@

\MKin{now is time to talk about events? or maybe right after presenting the rules}

\begin{center}
  \judgbox{\pstepto{\runtimetermvar}{\runtimetermvar[']}{\event}}{,,Primitive step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
  \typerule{$e-\text{get}-\in$}{
    \sandboxtag=\statevar.\sandboxtag &
    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
    \loc+n\in\statevar.\heap[^{\sandboxtag}]
  }{
    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{\heap[^{\sandboxtag}](\loc+n)}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
  }{e-get-in}
  %
  \typerule{$e-\text{get}-\notin$}{
    \sandboxtag=\statevar.\sandboxtag &
    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
    \loc+n\notin\statevar.\heap[^{\sandboxtag}]
  }{
    \pstepto{\rtt{\statevar}{\lget{x}{n}}}{\rtt{\statevar}{1729}}{\tup{\ev{Get}\ \loc\ n}{\sandboxtag}}
  }{e-get-notin}
  %
  \typerule{$e-\text{set}-\in$}{
    \sandboxtag=\statevar.\sandboxtag &
    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
    \loc+n\in\statevar.\heap[^{\sandboxtag}] &
    \heap[^{\sandboxtag}_1] = \seth{\statevar.\heap[^{\sandboxtag}]}{\loc+n}{\valueexpr}
  }{
    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}_1]}{\heap[^{\sandboxtag}]}}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
  }{e-set-in}
  %
  \typerule{$e-\text{set}-\notin$}{
    \sandboxtag=\statevar.\sandboxtag &
    \statevar.\store(x)=\tup{\loc}{\sandboxtag,\poison,m} \\
    \loc+n\notin\statevar.\heap[^{\sandboxtag}]
  }{
    \pstepto{\rtt{\statevar}{\lset{x}{n}{\valueexpr}}}{\rtt{\statevar}{\valueexpr}}{\tup{\ev{Set}\ \loc\ n\ \valueexpr}{\sandboxtag}}
  }{e-set-notin}
  %
  \typerule{$e-\text{new}$}{
    \fresh{\statevar}{z} &
    \fresh{\statevar}{\loc} &
    \heap[^{\sandboxtag}_1]=\growh{\statevar.\heap[^{\sandboxtag}]}{n} &
    \store[_1]=\storeel{z}{\loc}{\statevar.\sandboxtag}{\poisonless}{n},\statevar.\store
  }{
    \pstepto{\rtt{\statevar}{\lnew{x}{n}{\expr}}}{\rtt{\statevar\subst{\heap[^{\sandboxtag}]}{\heap[^{\sandboxtag}_1]}\subst{\store}{\store[_1]}}{\expr\subst{z}{x}}}{\tup{\ev{Alloc}\ \loc\ n}{\sandboxtag}}
  }{e-new}
  %
  \typerule{$e-\text{dealloc}$}{
    \statevar.\store(x) = \tup{\loc}{\statevar.\sandboxtag,\poison,n} &
    \store[_1]=\seth{\statevar.\store}{x}{\tup{\loc}{\statevar.\sandboxtag,\poisoned,n}}
  }{
    \pstepto{\rtt{\statevar}{\ldelete{x}}}{\rtt{\statevar\subst{\store}{\store[_1]}}{0}}{\tup{\ev{Dealloc}\ \loc}{\sandboxtag}}
  }{e-dealloc}
  %
  \typerule{$e-\text{abort}$}{
  }{
    \pstepto{\rtt{\statevar}{\labort}}{\lightning}{\lightning}
  }{e-abort}
\end{center}

Reading the value of a pointer is done by looking up the metadata in the store (\Cref{tr:e-get-in,tr:e-get-notin}).
This is done without any bounds-checking and always emits a $\ev{Get}\ \loc\ n$ event regardless of whether the access is out of bounds or not.
Similarily when setting a value (\Cref{tr:e-set-in,tr:e-set-notin}), a $\ev{Set}\ \loc\ n\ \valueexpr$ event is emitted regardless of wether the set is in or out of bounds.
Allocation of objects (\Cref{tr:e-new}) does the expected, however, it makes use of the judgement $\fresh{\cdot}{\cdot}$ which is used throughout this paper to explicitly mark when and with respect to what environment a fresh variable is introduced.
Deallocation (\Cref{tr:e-dealloc}) simply sets the poison--tag in the metadata of the associated pointer to be poisoned ($\poisoned$).
The only way to abort an execution (\Cref{tr:e-abort}) is with an explicit call to $\labort$.
However, evaluation of programs can get stuck, e.g., if $x$ is not bound in $\ldelete{x}$.

In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.

\begin{center}
  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
  %
  \typerule{$e-\text{ectx}$}{
    \pstepto{\rtt{\statevar}{\expr}}{\rtt{\statevar[']}{\expr[']}}{\event}
  }{
    \estepto[]{\rtt{\statevar}{\ectx\hole{\expr}}}{\rtt{\statevar[']}{\ectx\hole{\expr[']}}}{\event}
  }{e-ectx}
  %
  \typerule{$e-\text{stuck}$}{
    \pstepto{\rtt{\statevar}{\expr}}{\lightning}{\lightning}
  }{
    \estepto[]{\rtt{\statevar}{\ectx\hole{\expr}}}{\lightning}{\lightning}
  }{e-stuck}
  %
  \typerule{$e-\text{ret}$}{
    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
    \commlib\vdash\text{foo}:\Omega.\sandboxtag &
    \operatorname{comm}(\sandboxtag)=\comm
  }{
    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
  }{e-ret-notsame}
  %
  \typerule{$e-\text{start}$}{
    \lfunction{\text{main}}{x}{e} \in \statevar.\library
  }{
    \estepto[]{\rtt{\statevar}{\lcall{\text{main}}{0}}}{\rtt{\statevar\subst{\ctx}{\sandboxtag}\subst{\tup{\hole{\cdot}}{\text{main}}}{\ectxstack}}{\expr\subst{0}{x}}}{\tup{\ev{Start}}{\comp}}
  }{e-start}
  %
  \typerule{$e-\text{call}-\text{notsame}$}{
    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
    \commlib\vdash\text{foo}:\Omega.\sandboxtag &
    \operatorname{comm}(\sandboxtag)=\comm
  }{
    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
  }{e-call-notsame}
  %
  \judgbox{\commlib\vdash\text{foo}:\sandboxtag}{,,Given execution is in $\sandboxtag$, the symbol associated with the name $\text{foo}$ is in $\neg\sandboxtag$.''}
  %
  \typerule{ctx-to-comp}{
    \text{foo}\in\commlib
  }{
    \commlib\vdash\text{foo}:\ctx
  }{ctx-to-comp}
  %
  \typerule{comp-to-ctx}{
    \text{foo}\notin\commlib
  }{
    \commlib\vdash\text{foo}:\comp
  }{comp-to-ctx}
\end{center}

Contextual steps are as expected (\Cref{tr:e-ectx,tr:e-stuck}) up to calling (\Cref{tr:e-call-notsame}) and returning.
For calls and returns, a bit of work has to be done to emit the right events.
For example, the call to the main function emits just $\ev{Start}$ (\Cref{tr:e-start}) instead of $\ev{Call}\ \comptoctx\ \text{main}\ 0$, which is a design choice this paper does for convenience when reasoning about a call-chain.
Ordinary calls, however, need to check wether a context switch is happening or not and push the current continuation and the name of the invoked function on the stack in the state (\Cref{tr:e-call-notsame}).
Context switches are checked by looking at the current control--tag and wether the involved function or surrounding context is part of the list of names of component--level symbols ($\commlib$) or not (\Cref{tr:ctx-to-comp,tr:comp-to-ctx}).

\begin{center}
  \judgbox{\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}}{,,Program $\prog{\library_{\ctx}}{\library_\comp}$ evaluates to $\runtimetermvar$, emitting trace $\trace$.''}
  %
  \typerule{ep-run}{
    \link{\library_{\ctx}}{\library_{\comp}}=\library &
    \commlib=\operatorname{dom}\library_{\comp} \\
    \estepto{\rtt{\tup{\library}{\commlib,\emptyset}}{\lcall{\text{main}}{0}}}{\rtt{\statevar}{\valueexpr}}{\trace}
  }{
    \progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\rtt{\statevar}{\valueexpr}}{\trace}
  }{ep-run}
  %
  \typerule{ep-crash}{
    \link{\library_{\ctx}}{\library_{\comp}}=\library &
    \commlib=\operatorname{dom}\library_{\comp} \\
    \estepton{\rtt{\tup{\library}{\commlib,\emptyset}}{\lcall{\text{main}}{0}}}{\lightning}{\trace\cdot\lightning}
  }{
    \progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\lightning}{\trace\cdot\lightning}
  }{ep-crash}
\end{center}

The notation $\tup{\library}{\commlib,\emptyset}$ unfolds to $\statevar=\tup{\cfstate}{\comp,\memstate}$, where $\cfstate=\tup{\library}{\commlib,\hole{\cdot}}$ and $\memstate=\tup{\hole{\cdot}}{\hole{\cdot},\hole{\cdot}}$.
Linking $\link{\library_{\ctx}}{\library_{\comp}}$, while gluing $\library_{\ctx}$ and $\library_{\comp}$ together, ensures that there is no symbol with the name $\text{main}$ in $\library_{\comp}$.
Programs can either be run successfully (\Cref{tr:ep-run}) or fail (\Cref{tr:ep-crash}), which requires the n-step relation instead of star step, since star step models a big--step style semantics without failure.
To make use of \Cref{def:propsat,def:proprsat}, {\em behavior} is introduced as follows.

\begin{definition}[Program Behavior]\label{def:behav}
  Given a program $\prog{\library_{\ctx}}{\library_{\comp}}$, the behavior of it is a set containing all traces $\trace$ such that there is a runtime--term $\runtimetermvar$ with $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$.
\end{definition}

>>>>

%\section{Case Study: Memory Safe and Cryptographic Constant Time Code}\label{sec:casestud}
\section{Case Study: Preliminary Definitions\pages{3}}\label{sec:casestud:defs}

@@

\subsection{Shared Language Definitions}\label{subsec:cs:defs}

\begin{gather*}
  \begin{aligned}
  \mi{(Expressions)}~\expr&\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr} \mid \lnew{\expr[_{1}]}{\expr[_{2}]} \mid \ldelete{x} \mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \\
    & \mid \llet{x}{\expr[_{1}]}{\expr[_{2}]}  \mid \lreturn{\expr} \mid \lcall{g}{\expr} \mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort \\
  \mi{(Symbols)}~&\function\bnfdef\ \lfunction{g}{x}{e}\hspace{0.5cm}
  \mi{(Libraries)}~\library\bnfdef\ \hole{\cdot} \mid \function,\library \\
  \mi{(Programs)}~&\prog{\library_{\ctx}}{\library_{\comp}}\hspace{0.5cm}
  \mi{(Substitutions)}~\substvar\bnfdef \hole{\cdot} \mid \subst{\valueexpr}{x},\gamma \\
  \end{aligned}
\end{gather*}

Above is a fraction of the shared syntax of all the programming languages of this paper.
The syntax and semantics are mostly standard.
However, when allocating pointers with $\lnew{\expr[_{1}]}{\expr[_{2}]}$, the expression $\expr[_{1}]$ determines the initial value of the objects and $\expr[_{2}]$ the overall size of an array.
The data associated to pointers resides in the dynamic state $\Omega$ that contains sandboxed heaps, one for the context and the other for the component.
Because of the sandboxing, locations pointing into heaps must be associated with a control--tag to know the heap they point into.
Other metadata in $\Omega$ entails a list of callable objects with information on wether they belong to the context or the component.
That way, when calling a callable object, the semantics can take care of emitting appropriate events that signal a context switch.
For the languages considered in this paper, no callable objects are first--class.
So, the only values are pairs, pointers, and natural numbers.
Another non--standard part of the languages is the concept of the poison--tag.
The poison--tag marks a location as freed ($\poisoned$) without removing the information associated to the freed pointer from $\Omega$.
This measure prevents the semantics from getting stuck when an already freed pointer is freed again.
Because of this, it is possible that programs like $\llet{x}{\lnew{0}{5}}{\ldelete{x};\ldelete{x}}$ reduce and emit a trace along the lines of $\specificev{Alloc\ \loc\ \sandboxtag\ 5}\cdot\specificev{Dealloc\ \loc\ \sandboxtag}\cdot\specificev{Dealloc\ \loc\ \sandboxtag}$.

@@

Possible events that are emitted during execution are similar to the ones of the specification trace model.
The key difference is that, as standard in secure compilation work, the traces have a call and return event that signals context switches.
Hereby, a $\specificev{Call\ \ctxtocomp\ foo\ \valueexpr}$ and $\specificev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from context-- to component--level.
Contrary, $\specificev{Call\ \comptoctx\ foo\ \valueexpr}$ and $\specificev{Return\ \ctxtocomp\ \valueexpr}$ signal that program execution transitions from component-- to context--level.
The environmental semantics emits the tag $\nocomm$ for calls from context-- to context--level or component-- to component--level.
In the following, $\neg\ctx=\comp$ and $\neg\comp=\ctx$ and also $\operatorname{comm}(\ctx)=\ \ctxtocomp$ and $\operatorname{comm}(\comp)=\ \comptoctx$.

\begin{center}
  \judgbox{\estepto[]{\runtimetermvar}{\runtimetermvar}{\event}}{,,Contextual step from runtime--term $\runtimetermvar$ to $\runtimetermvar[']$ emitting event $\event$.''}
  %
  \typerule{$e-\text{ret}$}{
    \statevar.\ectxstack=\tup{\ectx}{\text{foo}},\ectxstack['] &
    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
  }{
    \estepto[]{\rtt{\statevar}{\ectx[']\hole{\lreturn{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\ectxstack[']}{\ectxstack}}{\ectx\hole{\valueexpr}}}{\tup{\ev{Ret}\ \comm\ \valueexpr}{\statevar.\sandboxtag}}
  }{e-ret-notsame}
  %
  \typerule{$e-\text{call}-\text{notsame}$}{
    \lfunction{\text{foo}}{x}{e} \in \statevar.\library &
    \commlib\vdash\text{foo}:\Omega.\sandboxtag\dashv\comm
  }{
    \estepto[]{\rtt{\statevar}{\ectx\hole{\lcall{\text{foo}}{\valueexpr}}}}{\rtt{\statevar\subst{\neg\statevar.\sandboxtag}{\sandboxtag}\subst{\tup{\ectx}{\text{foo}},\statevar.\ectxstack}{\ectxstack}}{\expr\subst{\valueexpr}{x}}}{\tup{\ev{Call}\ \comm\ \text{foo}\ \valueexpr}{\statevar.\sandboxtag}}
  }{e-call-notsame}
\end{center}

The environmental semantics is also mostly straightforward.
In \Cref{tr:e-ret-notsame,tr:e-call-notsame} the judgement $\commlib\vdash\text{foo}:\Omega.\sandboxtag$ checks wether $\text{foo}$ is a component--level name by looking it up in the list of component-level names $\commlib$ and emits the appropriate transfer tag, i.e., either $\comptoctx$ or $\ctxtocomp$.
Additional rules that are left out ensure that, e.g., when calling the $\text{main}$ function, the event $\ev{Start}$ is emitted, which is a design choice this paper does for convenience when reasoning about call--chains.

The top--level execution $\progstepto{\prog{\library_{\ctx}}{\library_{\comp}}}{\runtimetermvar}{\trace}$ constructs an initial state $\Omega$ by linking $\library_{\ctx}$ and $\library_{\ctx}$ and then starts execution by calling the $\text{main}$ function.
The trace $\trace$ emitted during that execution serves as abstraction of the behavior of the program enabling the use of \Cref{def:propsat,def:proprsat}.
